Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-04-28 overleaf/compiler.tex
Edits to Section 4.

\begin{enumerate}
  \item
        \textbf{Parsing}:
        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
  \item
        \textbf{Syntax checking}:
        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
  \item
        \textbf{3-address CFG conversion}:
        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}

        Convert the restricted-syntax AST to a three-address control-flow graph.
        To do this, first,
        add an empty basic block to the CFG and mark it as current.
        Next, for each statement in the restricted AST's function body,
        process the statement.
        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
        Rules for processing each kind of statement are given below:
        \begin{enumerate}
          \item
                \textbf{For-loops}:
                Create new basic blocks for
                the loop condition
                (the \emph{condition-block}),
                the loop body
                (the \emph{body-block}),
                and the code after the loop
                (the \emph{after-block}).
                Insert a jump from the end of the current block to the condition-block.
                Then, mark the condition-block as the current block.
                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
                an edge from the current block to the
                body-block labeled ``TRUE''.
                Then, set the body-block to be the current block
                and process all statements in the AST's loop body.
                Finally, insert a jump to the condition-block and set the after-block as current.
          \item
                \textbf{If-statements}:
                Create new basic blocks for
                the ``then'' statements of the if-statement
                (the \emph{then-block}),
                the ``else'' statements of the if-statement
                (the \emph{else-block}),
                and the code after the if-statement
                (the \emph{after-block}).
                At the end of the current block,
                insert a conditional jump to the then-block or else-block
                depending on the if-statement condition in the AST.
                Next,
                mark the then-block as current,
                process all then-statements,
                and add a jump to the after-block.
                Similarly,
                mark the else-block as current,
                process all else-statements,
                and add a jump to the after-block.
                Finally, set the after-block to be the current block,
                and give it a \emph{merge condition} property equal to the condition of the if-statement.
          \item
                \textbf{Assignments}:
                In the restricted-syntax AST,
                the left-hand side of assignments
                can be a variable or an array subscript.
                If it is an array subscript, e.g., {\sf A[i] = x},
                change the statement to {\sf A = Update(A, i, x)}.
                If the statement is not already three-address code,
                for each sub-expression in the right-hand side of the assignment,
                insert an assignment to a temporary variable.
        \end{enumerate}
  \item
        \textbf{SSA conversion}:
        Convert the 3-address CFG to SSA with Cytron's algorithm.
\end{enumerate}

@@

Once the compiler converts the code to SSA,
it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
From the 3-address CFG conversion step,
$\phi$-nodes corresponding to if-statements will be in a basic block
with the merge condition property.
For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.

@@

Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes
and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated
during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.

@@

MOTION supports FOR loops and SIMD operations, so translation from MPC source to MOTION C++ code is relatively straightforward.
% Variable naming
Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g. {\sf sum!2} would be translated to {\sf sum\_2}). Plaintext variables follow the same naming convention as shared variables but are prefixed with \texttt{\_MPC\_PLAINTEXT\_}.  The shared representation of constants are named \texttt{\_MPC\_CONSTANT\_} followed by the literal constant (e.g. the shared constant \texttt{0} would be named \texttt{\_MPC\_CONSTANT\_0}).

% Function preamble
The generated MOTION code begins with the declaration of all variables used in the function, including loop counters.  If a variable is a vectorized array, it is initialized to a correctly-sized array of empty MOTION shares.  Additionally, each plaintext variable and parameter has a shared counterpart declared.  Next, all constant values which are used as part of shared expressions are initialized as a shared input from party 0.  Finally, all plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.

% Non-vectorized assignments and returns
Once the function preamble is complete, the MPC source is translated into C++ one statement at a time.  If there is no vectorization present in a statement, translation to C++ is straightforward: outside of Mux statements and array updates, non-vectorized assignments, expressions, and returns directly translated into their C++ equivalents.  Unvectorized Mux statements are converted to MOTION's Mux member function on the condition variable.  Array updates are translated into two C++ assignments: one to update the value in the original array and one to assign the new array.

@@

% Phi nodes and FOR loops
Phi nodes are broken into two components: the ``false'' branch which assigns the initial value of the phi node and the ``true'' branch which assigns the phi node's backedge.  The assignment of the false branch occurs right before the Phi node's enclosing loop.  Inside of the Phi node's enclosing loop, a C++ \texttt{if} statement is inserted to only assign the true branch of the phi node after the first iteration.  MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.  The loop counter is initialized before any phi nodes in the loop as they could rely on it; there is no other special handling of FOR loops.

\begin{table*}
\begin{tabular}{cc}
\begin{minipage}{0.5\textwidth}
{\small
\begin{pythonn}
for i in range(N):
   tmp = PHI(arr[i], val!0)
	 ...
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.5\textwidth}
{\small
\begin{cppp}
_MPC_PLAINTEXT_i = 0;
tmp = arr[_MPC_PLAINTEXT_i];
for (; _MPC_PLAINTEXT_i < _MPC_PLAINTEXT_N; _MPC_PLAINTEXT_i++) {
   if (_MPC_PLAINTEXT_i != 0) {
      tmp = val_0;
	 }
	 ...
}
\end{cppp}
}
\end{minipage}

@@

% SIMD representation
Vectorization is handled with utility functions to manage accessing and updating slices of arrays.  All SIMD values are stored in non-vectorized form as 1-dimensional \texttt{std::vector}s in row-major order.  Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the unvectorized representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.  Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.  Therefore the translation of an expression containing SIMD values is identical to that of expressions without SIMD values once vectorized accesses are inserted.

Similarly, the \texttt{vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.  This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the non-vectorized representation.
\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.  SIMD values are never directly assigned to non-SIMD values.

\begin{table*}
\begin{tabular}{cc}
\begin{minipage}{0.5\textwidth}
{\small
\begin{pythonn}
sum!4[I] = ADD_SIMD(sum!3[I], p[I, j])
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.5\textwidth}
{\small
% Ben: This line is too long but I don't know how to break it well
\begin{cppp}
vectorized_assign(sum_4, {_MPC_PLAINTEXT_N}, {true}, {}, vectorized_access(sum_3, {_MPC_PLAINTEXT_N}, {true}, {}) + vectorized_access(p, {_MPC_PLAINTEXT_N, _MPC_PLAINTEXT_D}, {true, false}, {_MPC_PLAINTEXT_j}));
\end{cppp}
}
\end{minipage}

@@

% vectorized_update()
\ben{Technically this is semantically the same as non-vectorized update but we use a utility function to combine the steps into a single C++ statement - should it still be included?}
Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.  Instead of separating the array update from the assignment of the new array, these steps are combined with the \texttt{vectorized\_update()} utility function.  This function operates identically to \texttt{vectorized\_assign()}, however it additionally returns the array after the assignment occurs.  This value is then used for the assignment to the new variable.

% raise_dim() implementation
\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...}
Raising the dimensions of an expression uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.  The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.  There is also a special case of the \texttt{lift()} function which occurs when we are raising an array.  This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.  In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).

% Ben (note as comment since tables aren't fixed placement): I only include the translation of the expression itself (instead of an assignment) since the assignment includes the extra junk from a vectorized_assign.  Also none of the raise_dims in biometric use the dimension index which I think is the most interesting/important part of the lift expression translation so I made up a toy example.
\begin{table*}
\begin{tabular}{cc}
\begin{minipage}{0.5\textwidth}
{\small
\begin{pythonn}
raise_dim(i + j, (i:N, j:M))
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.5\textwidth}
{\small
\begin{cppp}
lift(std::function([&](const std::vector<std::uint32_t> &idxs){return idxs[0] + idxs[1];}), {_MPC_PLAINTEXT_N, _MPX_PLAINTEXT_M})
\end{cppp}
}
\end{minipage}

@@

% drop_dim() implementation
Dropping dimensions use the \texttt{drop\_dim()} and \texttt{drop\_dim\_monoreturn()} utility functions.  They function identically but the latter returns a scalar for the case when the final dimension of an array is dropped.  These functions take the non-vectorized representation of an array, along with the dimensions of that array, and return the array with the final dimension dropped.
\ben{drop\_dim() taking the true branch of phi node values doesn't need to be covered here, right?}

% Shared/plaintext conversions
\ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols \ben{Should that go here?}.  MOTION does not support publicly-known constants for these protocols, so all conversions from plaintext values to shares are performed by providing the plaintext value as a shared input from party 0.  Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions into a plaintext value.  This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.  Since variables are often initialized to a common constant value (e.g. 0), this approach decreases the number of input gates by only creating a shared input for each constant.  Loop counters must still be converted to a shared value on each iteration that they are used, however we only output this conversion when necessary to prevent loops where the counter is not used in a shared expression from increasing the number of input gates.

% Tons of copies != runtime costs
Due to the SSA translation phase as well as the various conversions to and from SIMD values which our utility functions perform, our generated MOTION code often includes many copies of arrays and scalar values.  These copies do not incur a runtime cost as the arrays simply hold pointers to the underlying shares, so no new shares or gates are created as a result of this copying. \ben{Is there anything else to say here?  Should this be placed somewhere else?}

>>>>

%\begin{enumerate}
%  \item
        \paragraph{Parsing:}
        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
%  \item
        \paragraph{Syntax checking:}
        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
%  \item
        \paragraph{3-address CFG conversion:}
        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}

        Convert the restricted-syntax AST to a three-address control-flow graph.
        To do this, first,
        add an empty basic block to the CFG and mark it as current.
        Next, for each statement in the restricted AST's function body,
        process the statement.
        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
        Rules for processing each kind of statement are given below:
        \begin{enumerate}
          \item
                \textbf{For-loops}:
                Create new basic blocks for
                the loop condition
                (the \emph{condition-block}),
                the loop body
                (the \emph{body-block}),
                and the code after the loop
                (the \emph{after-block}).
                Insert a jump from the end of the current block to the condition-block.
                Then, mark the condition-block as the current block.
                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
                an edge from the current block to the
                body-block labeled ``TRUE''.
                Then, set the body-block to be the current block
                and process all statements in the AST's loop body.
                Finally, insert a jump to the condition-block and set the after-block as current.
          \item
                \textbf{If-statements}:
                Create new basic blocks for
                the ``then'' statements of the if-statement
                (the \emph{then-block}),
                the ``else'' statements of the if-statement
                (the \emph{else-block}),
                and the code after the if-statement
                (the \emph{after-block}).
                At the end of the current block,
                insert a conditional jump to the then-block or else-block
                depending on the if-statement condition in the AST.
                Next,
                mark the then-block as current,
                process all then-statements,
                and add a jump to the after-block.
                Similarly,
                mark the else-block as current,
                process all else-statements,
                and add a jump to the after-block.
                Finally, set the after-block to be the current block,
                and give it a \emph{merge condition} property equal to the condition of the if-statement.
          \item
                \textbf{Assignments}:
                In the restricted-syntax AST,
                the left-hand side of assignments
                can be a variable or an array subscript.
                If it is an array subscript, e.g., {\sf A[i] = x},
                change the statement to {\sf A = Update(A, i, x)}.
                If the statement is not already three-address code,
                for each sub-expression in the right-hand side of the assignment,
                insert an assignment to a temporary variable.
        \end{enumerate}
  %\item
        \paragraph{SSA conversion}:
        Convert the 3-address CFG to SSA with Cytron's algorithm.
%\end{enumerate}

@@

Once the compiler converts the code to SSA,
it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
From the 3-address CFG conversion step,
$\phi$-nodes corresponding to if-statements will be in a basic block
with the merge condition property.
For example, if {\sf X!3 = $\phi$(X!1,X!2)} is in a block with merge condition {\sf C},
the compiler transforms it into {\sf X!3 = MUX(C, X!1, X!2)}.
Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.

@@

Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes
and we write {\sf PHI} in MPC Source. A pseudo $\phi$-node {\sf X!1 = PHI(X!0,X!2)} in a loop header is evaluated
during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to {\sf X!0}, otherwise, it evaluates to {\sf X!2}.

@@

MOTION supports FOR loops and SIMD operations, so translation from MPC source to MOTION C++ code is relatively straightforward.
% Variable naming
\paragraph{Variable declarations:}
Our compiler performs taint analysis on MPC Source based on the type system in~\secref{sec:syntax}. The analysis is described in the following section;
for the most part, it is a standard taint analysis that propagates that \texttt{shared} qualifiers throughout the routine. 
Taint analysis assigns \texttt{shared} or \texttt{plain} qualifiers to all local variables, which is important for code generation.
Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g. {\sf sum!2} would be translated to {\sf sum\_2}). Plaintext variables follow the same naming convention as shared variables but are prefixed with {\sf \_MPC\_PLAINTEXT\_}.  The shared representation of constants are named {\sf \_MPC\_CONSTANT\_} followed by the literal constant (e.g. the shared constant 0 would be named {\sf \_MPC\_CONSTANT\_0}).

% Function preamble
The generated MOTION code begins with the declaration of all variables used in the function, including loop counters.  If a variable is a vectorized array, it is initialized to a correctly-sized array of empty MOTION shares.  Additionally, each plaintext variable and parameter has a shared counterpart declared.  Next, all constant values which are used as part of shared expressions are initialized as a shared input from party 0.  Finally, plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts. (Importantly, as we mention, they are converted only if they are needed as an argument to a shared operation; one can think of it as an upcast from \texttt{plain} to \texttt{share}.)

% Non-vectorized assignments and returns
\paragraph{Code generation:}
Once the function preamble is complete, the MPC Source is translated into C++ one statement at a time. The linear structure of MPC Source enables translation. If there is no vectorization present in a statement, translation to C++ is straightforward: outside of MUX statements and array updates, non-vectorized assignments, expressions, and returns directly translate into their C++ equivalents.  Non-vectorized MUX statements are converted to MOTION's MUX member function on the condition variable.  Array updates are translated into two C++ assignments: one to update the value in the original array and one to assign the new array as shown in Listing~\ref{tab:motion_translation_array_updates}.

@@

% Phi nodes and FOR loops
Pseudo PHI nodes are broken into two components: the ``FALSE'' branch which assigns the initial value of the PHI node and the ``TRUE'' branch which assigns the PHI node's back-edge.  The assignment of the false branch occurs right before the PHI node's enclosing loop.  Inside of the PHI node's enclosing loop, a C++ \texttt{if} statement is inserted to only assign the true branch of the PHI node after the first iteration. MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.  The loop counter is initialized before any PHI nodes in the loop as they could rely on it; there is no other special handling of FOR loops. \ana{Add the reference to the corresponding figure somewhere in this paragraph.}

\begin{table*}
\begin{tabular}{cc}
\begin{minipage}{0.33\textwidth}
{\small
\begin{pythonn}
for i in range(N):
   tmp = PHI(arr[i], val!0)
	 ...
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.66\textwidth}
{\small
\begin{cppp}
_MPC_PLAINTEXT_i = 0;
tmp = arr[_MPC_PLAINTEXT_i];
for (; _MPC_PLAINTEXT_i < _MPC_PLAINTEXT_N; _MPC_PLAINTEXT_i++) {
   if (_MPC_PLAINTEXT_i != 0) {
      tmp = val_0;
	 }
	 ...
}
\end{cppp}
}
\end{minipage}

@@

% SIMD representation
Vectorization is handled with utility functions to manage accessing and updating slices of arrays.  All SIMD values are stored in non-vectorized form as 1-dimensional \texttt{std::vector}s in row-major order. \ana{Does MOTION support just 1-dim SIMD arrays? If yes, mention it here.} Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value. \ana{Add the reference to the ADD example figure here.} Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression. \ana{Is it correct to say that the operators are overloaded? If yes, use that term?} Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.

Similarly, the \texttt{vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.  This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the non-vectorized representation.
\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.  SIMD values are never directly assigned to non-SIMD values. \ana{Yes, this is implementation-specific, you can remove.}

\begin{table*}
\begin{tabular}{cc}
\begin{minipage}{0.33\textwidth}
{\small
\begin{pythonn}
sum!4[I] = ADD_SIMD(sum!3[I], p[I, j])
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.66\textwidth}
{\small
% Ben: This line is too long but I don't know how to break it well
\begin{cppp}
vectorized_assign(sum_4, {_MPC_PLAINTEXT_N}, {true}, {}, 
      vectorized_access(sum_3, {_MPC_PLAINTEXT_N}, {true}, {}) + 
      vectorized_access(p, {_MPC_PLAINTEXT_N, _MPC_PLAINTEXT_D}, {true, false}, 
           {_MPC_PLAINTEXT_j}));
\end{cppp}
}
\end{minipage}

@@

% vectorized_update()
\ben{Technically this is semantically the same as non-vectorized update but we use a utility function to combine the steps into a single C++ statement - should it still be included?}
\ana{I would leave this in the TR but it will go from the paper. Lots of stuff in Secs. 4 and 5 will go from the paper.} 
Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.  Instead of separating the array update from the assignment of the new array, these steps are combined with the \texttt{vectorized\_update()} utility function.  This function operates identically to \texttt{vectorized\_assign()}, however it additionally returns the array after the assignment occurs.  This value is then used for the assignment to the new variable.

% raise_dim() implementation

\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{Yes, that's true. To make matters worse, most of the explanation on this comes in the next section. But I'll leave as is for now.}
\paragraph{Reshaping and raising dimensions:}
Raising the dimensions of a scalar or array uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.  The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.  There is also a special case of the \texttt{lift()} function which occurs when we are raising an array.  This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.  In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is). \ana{Add the ref. to the figure. Also, try to rewrite. I understand what the paragraph is saying but it is too technical and does not convey the higher-level goal of these. Try first to state what the goal is, and then talk about the details such as the lambda funcs.}

% Ben (note as comment since tables aren't fixed placement): I only include the translation of the expression itself (instead of an assignment) since the assignment includes the extra junk from a vectorized_assign.  Also none of the raise_dims in biometric use the dimension index which I think is the most interesting/important part of the lift expression translation so I made up a toy example.
\begin{table*}
\begin{tabular}{cc}
\begin{minipage}{0.33\textwidth}
{\small
\begin{pythonn}
raise_dim(i + j, (i:N, j:M))
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.66\textwidth}
{\small
\begin{cppp}
lift(std::function([&](const std::vector<std::uint32_t > &idxs) {return idxs[0] + idxs[1];}), 
      {_MPC_PLAINTEXT_N, _MPX_PLAINTEXT_M})
\end{cppp}
}
\end{minipage}

@@

% drop_dim() implementation
\paragraph{Dropping dimensions:}
Dropping dimensions use the \texttt{drop\_dim()} and \texttt{drop\_dim\_monoreturn()} utility functions.  They function identically but the latter returns a scalar for the case when the final dimension of an array is dropped.  These functions take the non-vectorized representation of an array, along with the dimensions of that array, and return the array with the final dimension dropped.
\ben{drop\_dim() taking the true branch of phi node values doesn't need to be covered here, right?} \ana{Right, I wouldn't add it here.}

% Shared/plaintext conversions
\ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols \ben{Should that go here?} \ana{Yes, I think it's good}.  MOTION does not support publicly-known constants for these protocols, so all conversions from plaintext values to shares are performed by providing the plaintext value as a shared input from party 0.  Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions into a plaintext value. \ana{Typo: conversions from plaintext values, right?} This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.  Since variables are often initialized to a common constant value (e.g. 0), this approach decreases the number of input gates by only creating a shared input for each initialization constant.  Loop counters must still be converted to a shared value on each iteration that they are used, however we only generate this conversion when necessary, i.e., when the counter flows to a shared computation. This is to prevent unnecessary increase in the number of input gates.

% Tons of copies != runtime costs
Due to the SSA translation phase as well as the various conversions to and from SIMD values which our utility functions perform, our generated MOTION code often includes many copies of arrays and scalar values.  These copies do not incur a runtime cost as the arrays simply hold pointers to the underlying shares, so no new shares or gates are created as a result of this copying. Cost in MPC programs is dominated by shares and computation on shares. \ben{Is there anything else to say here?  Should this be placed somewhere else?} \ana{It is clear to me! Will stay in the TR but likely will be cut from the paper.}
